---
title: "biostat776 - Project 1 - Solution"
author: "Zhengting (Johnathan) He"
date: "2021/9/9"
output: html_document
---

```{r "setup", include = FALSE}
require("knitr")
opts_knit$set(root.dir = "D:/OneDrive - Johns Hopkins/Course/140.776.71 - Statistical Computing/project/Project 1")
```

```{r}
# Set working directory and load packages
setwd("D:/OneDrive - Johns Hopkins/Course/140.776.71 - Statistical Computing/project/Project 1/biostats776-Project1")

require(tidyverse)
```

```{r}
# Import dataset
tuesdata <- tidytuesdayR::tt_load('2020-01-21')

spotify_songs <- tuesdata$spotify_songs
glimpse(spotify_songs)
```

# Part 1: Explore data

In this part, use functions from `dplyr` to answer the following questions. 

1. How many songs are in each genre? 

```{r}
spotify_songs %>%
    group_by(playlist_genre) %>%
    count() %>%
    knitr::kable()
```


2. What is average value of `energy` and `acousticness` in the `latin` genre in this dataset?

```{r}
spotify_songs %>%
    filter(playlist_genre == "latin") %>%
    summarise(energy_mean = mean(energy, na.rm = FALSE),
    acousticness_mean = mean(acousticness, na.rm = FALSE)) %>%
    knitr::kable()
```


3. Calculate the average duration of song (in minutes) across all subgenres. Which subgenre has the longest song on average?

```{r}
spotify_songs %>%
    mutate(duration_m = (duration_ms/1000)/60) %>%
    group_by(playlist_subgenre) %>%
    summarise(duration_mean = mean(duration_m, na.rm = FALSE)) %>%
    arrange(duration_mean) %>%
    knitr::kable()
```

```
# Subgenre "progressive electro house" has the longest song on average.
```


4. Make two boxplots side-by-side of the `danceability` of songs stratifying by whether a song has a fast or slow tempo. Define fast tempo as any song that has a `tempo` above its median value. On average, which songs are more danceable?

**Hint**: You may find the `case_when()` function useful in this part, which can be used to map values from one variable to different values in a new variable (when used in a `mutate()` call).

```{r}
spotify_songs %>%
    mutate(tempo_cat = case_when(tempo > median(tempo, na.rm = TRUE) ~ "fast",
                                 TRUE ~ "slow")) %>%
    mutate(tempo_cat = factor(tempo_cat, levels = c("fast", "slow"))) %>%
    ggplot(aes(x = tempo_cat, y = danceability)) +
    stat_boxplot(geom = "errorbar", width = 0.2) +
    geom_boxplot(aes(fill = tempo_cat)) +
    labs(title = "Danceability of Songs Stratified by Tempo") +
    xlab("Tempo Categoty of Songs") +
    ylab("Danceability of Songs") +
    scale_x_discrete(labels = c("Fast", "Slow")) +
    scale_fill_manual(values = c("#439fd3", "#f6b128"),
                      labels = c("Fast", "Slow"),
                      guide = "none") +
    theme_bw()
```

```
# On average, songs having slow tempo are more danceable.
```


# Part 2: Convert nontidy data into tidy data

Use the functions in `dplyr`, `tidyr`, and `lubridate` to perform the following steps to the `spotify_songs` dataset: 

1. Select only unique distinct rows from the dataset based on the `track_name` and `track_artist` columns (**Hint** check out the `distinct()` function in `dplyr`). 
2. Add a new column called `year_released` listing just the year that the song was released. (**Hint** check out the `ymd()` function in `lubridate` R package. Also, if you get a warning message with "failed to parse", check out the `truncated` argument in the `ymd()` function.).
3. Keep only songs that were released in or after 1980. 
4. Add a new column with the duration of the song in minutes
5. For each `year_released`, calculate the mean of at least 6 of the audio features (e.g. danceability, energy, loudness, etc), or descriptors (e.g. tempo, duration in minutes, etc). (**Hint**: If all has gone well thus far, you should have a dataset with 41 rows and 7 columns). 
6. Convert this wide dataset into a long dataset with a new `feature` and `mean_score` column

It should look something like this: 

```
year_released   feature   mean_score
<dbl>           <chr>       <dbl>
1980	Danceability	0.5633676		
1980	Energy	0.7107647		
1980	Loudness	-8.5211765		
1980	Valence	0.6333235		
1980	Tempo	124.1458529		
1980	Duration	4.2853662		
1981	Danceability	0.5697258		
1981	Energy	0.6967581		
1981	Loudness	-8.8678065		
1981	Valence	0.6650968	
```

### Notes

* You may need to use functions outside these packages to obtain this result.

* Note that the functions in the `dplyr` and `tidyr` package expect table-like objects (data frames or tibbles) as their input. You can convert data to these objects using the `as_tibble()` function in the `tibble` package.

* Do not worry about the ordering of the rows or columns. Depending on whether you use `gather()` or `pivot_longer()`, the order of your output may differ from what is printed above. As long as the result is a tidy data set, that is sufficient.


```{r}
songs.summary <- spotify_songs %>%
                     distinct(track_name, track_artist, .keep_all = TRUE) %>%
                     mutate(track_album_release_date = lubridate::ymd(track_album_release_date, truncated = 2)) %>%
                     mutate(year_released = format(track_album_release_date, format = "%Y")) %>%
                     mutate(year_released = as.numeric(year_released)) %>%
                     filter(year_released >= 1980) %>%
                     mutate(duration_m = (duration_ms/1000)/60) %>%
                     group_by(year_released) %>%
                     summarise(Danceability = mean(danceability, na.rm = TRUE),
                               Energy = mean(energy, na.rm = TRUE),
                               Loudness = mean(loudness, na.rm = TRUE),
                               Tempo = mean(tempo, na.rm = TRUE),
                               Duration = mean(duration_m, na.rm = TRUE),
                               Valence = mean(valence, na.rm = TRUE)) %>%
                     pivot_longer(- year_released, names_to = "feature", values_to = "mean_score")
songs.summary %>% knitr::kable()
```


# Part 3: Data visualization

In this part of the project, we will continue to work with our now tidy song dataset from the previous part. 

### Tasks 

Use the functions in `ggplot2` package to make a scatter plot of the six (or more) `mean_score`s (y-axis) over time (x-axis). For full credit, your plot should include: 

1. An overall title for the plot and a subtitle summarizing key trends that you found. Also include a caption in the figure with your name. 
2. Both the observed points for the `mean_score`, but also a smoothed non-linear pattern of the trend
3. All six (or more) plots should be shown in the one figure 
4. There should be an informative x-axis and y-axis label

Consider playing around with the `theme()` function to make the figure shine, including playing with background colors, font, etc.

### Notes

* You may need to use functions outside these packages to obtain this result.

* Note that the functions in the `dplyr` and `tidyr` package expect table-like objects (data frames or tibbles) as their input. You can convert data to these objects using the `as_tibble()` function in the `tibble` package.

* Don't worry about the ordering of the rows or columns. Depending on whether you use `gather()` or `pivot_longer()`, the order of your output may differ from what is printed above. As long as the result is a tidy data set, that is sufficient.


```{r}
songs.summary %>%
    ggplot(aes(x = year_released, y = mean_score)) +
    geom_point(aes(colour = feature), size = 1) +
    geom_smooth(aes(colour = feature)) +
    facet_wrap(~ feature) +
    labs(x = "Year When Album Released",
         y = "Mean of Audio Features",
         title = "Scatter Plot of 6 Mean Scores of Audio Features Over Time",
         subtitle = "Trend of mean scores of audio features seems stable over time, except for fluctuation in Tempo",
         caption = "Plotted by Zhengting (Johnathan) He") +
    scale_colour_manual(values = c("#0ebeff", "#47cf73", "#ae63e4", "#fcd000", "#ff3c41", "#76daff"),
                        guide = "none") +
    theme(panel.background = element_rect(fill='transparent', color='black'),
          panel.border = element_rect(fill='transparent', color='transparent'),
          panel.grid = element_blank(),
          plot.title = element_text(size = 14, face = "bold", color = "black"),
          plot.subtitle = element_text(size = 10, face = "bold", color = "black"),
          plot.caption = element_text(size = 10, color = "black"),
          axis.title.x = element_text(size = 12, face = "bold", color = "black"),
          axis.title.y = element_text(size = 12, face = "bold", color = "black"),
          axis.text = element_text(size = 8, color = "black"),
          strip.background = element_rect(fill='transparent', color='black'),
          strip.text.x = element_text(size = 12, face = "bold", color = "black"))
```


# Part 4: Make the worst plot you can! 

This sounds a bit crazy I know, but I want this to be FUN! Instead of trying to make a "good" plot, I want you to explore your creative side and make a really awful data visualization in every way. :)

### Tasks

Using the `spotify_songs` dataset (and it does not have to be the tidy dataset that we created in Part 2, it can be anything from the original dataset): 

1. Make the absolute worst plot that you can. You need to customize it in **at least 7 ways** to make it awful. 
2. In your document, write 1 - 2 sentences about each different customization you added (using bullets -- i.e. there should be at least 7 bullet points each with 1-2 sentences), and how it could be useful for you when you want to make an awesome data visualization.  


```{r}
spotify_songs %>%
    ggplot(aes(x = playlist_genre)) +
    geom_jitter(aes(y = valence)) +
    geom_jitter(aes(y = tempo)) +
    geom_jitter(aes(y = key)) +
    geom_jitter(aes(y = speechiness)) +
    geom_jitter(aes(y = acousticness)) +
    geom_jitter(aes(y = liveness)) +
    theme(panel.background = element_rect(fill = "grey50", colour = NA),
          panel.grid = element_line(colour = "grey42"), 
          panel.grid.major = element_line(size = rel(0.5)), 
          panel.grid.minor = element_line(size = rel(0.25)),
          axis.title.x = element_text(size = 40, face = "bold", color = "black"),
          axis.title.y = element_text(size = 40, face = "bold", color = "black"),
          axis.text = element_text(size = 40, color = "black"))
```

**Different customization I added**

+ **Facet** of the plot: I did not use `ggplot2::facet_wrap()` to separate characteristics of the songs - *such as danceability, energy, key, etc.* - to different facets. Thus, the jitter points for each characteristics are massed up together, and hard to see the trend.

+ **Type** of the plot: I used `ggplot2::geom_jitter()` function to make the plot, which **mainly apply to smaller datasets**, and is not a good way to compare the differences between characteristics in different playlist genre group.

+ **Theme** for the plot: I change the arguments in the `ggplots::theme()` function based on the default value of ggplots theme `theme_dark()`, which **makes the panel a dark background**, and hard to distinguish the points from the panel.

+ **Font size** of axis text: The font size of the axis text is too large, causing x-axis text overlap between each other.

+ **Font size** of axis title: The font size of both x-axis and y-axis are too large, making the panel small to the readers, and makes it even harder to infer the differences between characteristics in different playlist genre group. Also, the y-axis only shows one characteristic - *danceability*.

+ **No title and caption** for the plot: The plot has no title or caption, making readers hard to understand what the plot is aiming to visualize.

+ **No file or color** for different playlist genre groups: File or color by playlist genre groups may make it easier for readers to make comparison between these different playlist genre.


# Part 5: Make my plot a better plot! 

The goal is to take my sad looking plot and make it better! If you'd like an [example](https://twitter.com/drmowinckels/status/1392136510468763652), here is a tweet I came across of someone who gave a talk about how to zhoosh up your ggplots. 

```{r}
spotify_songs %>% 
  ggplot(aes(y=track_popularity, x=playlist_subgenre, fill = playlist_genre)) + 
  geom_violin() +
  facet_wrap( ~ playlist_genre, scales = "free_x")
```
### Tasks 

1. You need to customize it in **at least 7 ways** to make it better. 
2. In your document, write 1 - 2 sentences about each different customization you added (using bullets -- i.e. there should be at least 7 bullet points each with 1-2 sentences), describing how you improved it.


```{r}
spotify_songs %>%
    select(playlist_genre, valence, tempo, key, speechiness:acousticness, liveness) %>%
    pivot_longer(- playlist_genre, names_to = "characteristic", values_to = "value") %>%
    mutate(value = log10(value)) %>%
    ggplot(aes(x = playlist_genre, y = value, fill = playlist_genre)) +
    stat_boxplot(geom = "errorbar", width = 0.2) +
    geom_boxplot() +
    facet_wrap(characteristic ~ .) +
    labs(x = "Playlist Genre",
         y = "Audio Features on log 10 scale",
         title = "Box and Wisker Plot of 6 Audio Features In Different Playlist Genre",
         caption = "Plotted by Zhengting (Johnathan) He") +
    scale_fill_manual(values = c("#0ebeff", "#47cf73", "#ae63e4", "#fcd000", "#ff3c41", "#76daff"),
                        guide = "none") +
    scale_x_discrete(labels = c("Edm", "Latin", "Pop", "R&B", "Rap", "Rock")) +
    theme(panel.background = element_rect(fill = "white", colour = NA),
          panel.border = element_rect(fill = NA, colour = "grey20"),
          panel.grid = element_line(colour = "grey92"), 
          panel.grid.minor = element_line(size = rel(0.5)),
          plot.title = element_text(size = 14, face = "bold", color = "black"),
          plot.caption = element_text(size = 10, color = "black"),
          axis.title.x = element_text(size = 12, face = "bold", color = "black"),
          axis.title.y = element_text(size = 12, face = "bold", color = "black"),
          axis.text = element_text(size = 8, color = "black"),
          strip.background = element_rect(fill = "grey85", colour = "grey20"),
          strip.text.x = element_text(size = 12, face = "bold", color = "black"))
```

**Different customization you added**

+ Transform dataset **from wide to long**: Using the `tidyr::pivot_longer()` function to gather all audio features into one column, and prepare for visualization using facets by these features.

+ **Facet** of the plot: Using `ggplot2::facet_wrap()` to separate characteristics of the songs - *such as danceability, energy, key, etc.* - to different facets, allowing for comparison between song characteristics.

+ **Fill the color** of the plot by playlist genre: Using the `fill` argument in the `aes()` function to fill the color of the plot by playlist genrem, thus allow for a clearer comparison between playlist genre.

+ **Type** of the plot: Using `ggplot2::geom_boxplot()` instead of `ggplot2::geom_jitter()`, and log transform the audio features, to allow for a clearer comparisons of audio features between playlist genre.

+ **Font size** of axis text: The font size of the axis text is changed to appropriate size. Also using`ggplots::scale_x_discrete()` to customize the x-axis text.

+ **Font size** of axis title: The font size of both x-axis and y-axis are changed to appropriate size, also change to "Audio Features on log 10 scale".

+ **Title and caption** for the plot: Add title and caption for the plot to let readers understand the plot purpose and author for the plot.